### 2020-05-19 doc/traits report/traits report.Rmd
full filled seed mass embryo

---
title: "Alpine traits report"
author: "Eduardo Fernández Pascual"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(tidyverse); library(here); library(FactoMineR); library(cowplot)
library(ggthemes); library(ggrepel); library(flextable); library(ggpubr)

read.csv(here("results", "database", "traits.csv")) %>%
  merge(read.csv(here("results", "database", "germination.csv")), by = "TPLName") %>% 
  merge(read.csv(here("data", "elevations", "regions.csv")), by = "Region", all.x = TRUE) %>%
  mutate(Macroregion = as.character(Macroregion),
         Macroregion = ifelse(is.na(Macroregion), "Europe", Macroregion),
         Macroregion = as.factor(Macroregion)) -> master

read.csv(here("results", "database", "traits.csv")) %>%
  select(TPLName, Seed.mass, Embryo) -> tt

```

```{r fig1, echo = FALSE, fig.cap = "Embryo by region and alpine character (brackets are SD)"}
master %>%
  group_by(TPLName, Macroregion, Alpine) %>%
  summarise(rate = mean(MGR, na.rm = TRUE), 
            syn = mean(SYN, na.rm = TRUE)) %>%
  merge(tt, by = "TPLName") %>%
  select(TPLName, Macroregion, Alpine, rate, syn, Seed.mass, Embryo) %>%
  gather(Trait, Value, rate:Embryo) %>%
  filter(is.finite(Value)) %>%
  filter(Trait == "Embryo") %>%
  group_by(Trait, Alpine, Macroregion) %>%
  summarise(y = mean(Value), SD = sd(Value)) %>%
  ggplot(aes(Alpine, y, ymin = y - SD, ymax = y + SD, fill = Alpine)) +
    theme_wsj() +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.4) + 
  facet_wrap(~ Macroregion, scales = "fixed") +
  geom_errorbar(aes(color = Alpine), width=.2,
                position=position_dodge(.9))
```

```{r fig2, echo = FALSE, fig.cap = "Seed mass by region and alpine character (brackets are SD)"}
master %>%
  group_by(TPLName, Macroregion, Alpine) %>%
  summarise(rate = mean(MGR, na.rm = TRUE), 
            syn = mean(SYN, na.rm = TRUE)) %>%
  merge(tt, by = "TPLName") %>%
  select(TPLName, Macroregion, Alpine, rate, syn, Seed.mass, Embryo) %>%
  gather(Trait, Value, rate:Embryo) %>%
  filter(is.finite(Value)) %>%
  filter(Trait == "Seed.mass") %>%
  group_by(Trait, Alpine, Macroregion) %>%
  summarise(y = mean(Value), SD = sd(Value)) %>%
  ggplot(aes(Alpine, y, ymin = y - SD, ymax = y + SD, fill = Alpine)) +
    theme_wsj() +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.4) + 
  facet_wrap(~ Macroregion, scales = "fixed") +
  geom_errorbar(aes(color = Alpine), width=.2,
                position=position_dodge(.9))
```

```{r fig3, echo = FALSE,fig.cap = "Mean germination rate by region and alpine character (brackets are SD)"}
master %>%
  group_by(TPLName, Macroregion, Alpine) %>%
  summarise(rate = mean(MGR, na.rm = TRUE), 
            syn = mean(SYN, na.rm = TRUE)) %>%
  merge(tt, by = "TPLName") %>%
  select(TPLName, Macroregion, Alpine, rate, syn, Seed.mass, Embryo) %>%
  gather(Trait, Value, rate:Embryo) %>%
  filter(is.finite(Value)) %>%
  filter(Trait == "rate") %>%
  group_by(Trait, Alpine, Macroregion) %>%
  summarise(y = mean(Value), SD = sd(Value)) %>%
  ggplot(aes(Alpine, y, ymin = y - SD, ymax = y + SD, fill = Alpine)) +
    theme_wsj() +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.4) + 
  facet_wrap(~ Macroregion, scales = "fixed") +
  geom_errorbar(aes(color = Alpine), width=.2,
                position=position_dodge(.9))
```

```{r fig4, echo = FALSE, fig.cap = "Germination synchrony by region and alpine character (brackets are SD)"}
master %>%
  group_by(TPLName, Macroregion, Alpine) %>%
  summarise(rate = mean(MGR, na.rm = TRUE), 
            syn = mean(SYN, na.rm = TRUE)) %>%
  merge(tt, by = "TPLName") %>%
  select(TPLName, Macroregion, Alpine, rate, syn, Seed.mass, Embryo) %>%
  gather(Trait, Value, rate:Embryo) %>%
  filter(is.finite(Value)) %>%
  filter(Trait == "syn") %>%
  group_by(Trait, Alpine, Macroregion) %>%
  summarise(y = mean(Value), SD = sd(Value)) %>%
  ggplot(aes(Alpine, y, ymin = y - SD, ymax = y + SD, fill = Alpine)) +
    theme_wsj() +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.4) + 
  facet_wrap(~ Macroregion, scales = "fixed") +
  geom_errorbar(aes(color = Alpine), width=.2,
                position=position_dodge(.9))
```

```{r fig5, echo = FALSE, fig.cap = "Proportion of full out of sown seeds by region and alpine character (brackets are binomial ci)"}
master %>%
  group_by(Macroregion, Alpine) %>%
  summarise(Sown = sum(Sown, na.rm = TRUE), 
            Germinable = sum(Germinable, na.rm = TRUE)) %>%
  filter(Germinable < Sown) %>% data.frame -> props

binom::binom.confint(props$Germinable, props$Sown, method = "wilson") %>% 
  cbind(props) %>%
  ggplot(aes(Alpine, mean, ymin = lower, ymax = upper, fill = Alpine)) +
    theme_wsj() +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.4) + 
  facet_wrap(~ Macroregion, scales = "fixed") +
  geom_errorbar(aes(color = Alpine), width=.2,
                position=position_dodge(.9))
```