### 2020-06-18 doc/mcmc report/mcm-report.Rmd
map figure

---
title: "MCMC report"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(tidyverse); library(here)

source(here("src", "analysis.R"))

```

# Manuscript figures

```{r echo = FALSE, fig.height = 5.5}
fig1
```

# Final germination proportions

```{r echo = FALSE, fig.height = 3}
fgFull
```

```{r echo = FALSE, fig.height = 3}
fgAlpine
```

```{r echo = FALSE, fig.height = 3}
fgRegion
```

```{r echo = FALSE, fig.height = 3}
fgDormancy
```

```{r echo = FALSE, fig.height = 3}
fgPlanttype
```
```{r echo = FALSE, fig.height = 3}
fgLifeform
```

```{r echo = FALSE, fig.height = 3}
fgLifespan
```

# Mean germination time

```{r echo = FALSE, fig.height = 3}
mgtFull
```

```{r echo = FALSE, fig.height = 3}
mgtAlpine
```

```{r echo = FALSE, fig.height = 3}
mgtRegion
```

```{r echo = FALSE, fig.height = 3}
mgtDormancy
```

```{r echo = FALSE, fig.height = 3}
mgtPlanttype
```
```{r echo = FALSE, fig.height = 3}
mgtLifeform
```

```{r echo = FALSE, fig.height = 3}
mgtLifespan
```

# Uncertainty

```{r echo = FALSE, fig.height = 3}
uncFull
```

```{r echo = FALSE, fig.height = 3}
uncAlpine
```

```{r echo = FALSE, fig.height = 3}
uncRegion
```

```{r echo = FALSE, fig.height = 3}
uncDormancy
```

```{r echo = FALSE, fig.height = 3}
uncPlanttype
```
```{r echo = FALSE, fig.height = 3}
uncLifeform
```

```{r echo = FALSE, fig.height = 3}
uncLifespan
```