### 2020-09-14 doc/cover-letter.Rmd
1st submission 🍋

---
title: "Cover letter"
author: "Eduardo Fernández Pascual"
date: "14/09/2020"
output: word_document
bibliography: refs.bib
csl: new-phytologist.csl
link-citations: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Dear Editorial Board,

I submit the manuscript *The seed germination spectrum of alpine plants: a global meta-analysis* to be considered for publication as a **full paper** in New Phytologist.

## What hypotheses or questions does this work address?

This is a global meta-analysis of primary germination data contributed by 12 research groups from 8 alpine regions in 4 continents. We used 9,799 records representing 62 seed plant families and 661 species to assess whether the germination ecological patterns detected by local studies are generalized at the global level.

## How does this work advance our current understanding of plant science?

We describe a global alpine germination syndrome, information valuable for all scientists working in alpine environments. Our results indicate a convergence of the seed germination spectrum in strict alpine species. We also found interactive effects of phylogeny and seed morphology that point to trade-offs shaping the seed ecological spectrum.

## Why is this work important and timely?

This is the first global evaluation, for a coherent habitat, of the seed germination spectrum and its relationship with other aspects of the seed ecological spectrum. This was one of the objectives in the research agenda for trait-based seed ecology recently published by New Phytologist [@RN4655].

## References