### 2020-11-07 doc/cover-letter.Rmd

---
title: "NPH-MS-2020-34477.R1 - Response to review"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output: pdf_document
bibliography: refs.bib
csl: new-phytologist.csl
link-citations: yes
always_allow_html: true
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Dear Editor,

Thank you for your positive comments and feedback. We submit a revised version of our manuscript, following your suggestions. In addition to the specific points commented below, we have revised the whole manuscript to improve its clarity.

> **1) Line 37: This is minor but I’m not in favour of using “meta-analyse” as a verb. Perhaps change to “We performed a meta-analysis…” I realize this uses a few more words but it reads better to my mind.**

We have done the suggested change.

> **2) Line 270: You imply (although it is a little vague) that for “some” species you filled in their values based on the average values for the genus or family. Please state how many species this was done for (and how many were filled in with genus versus family info) and state clearly whether this is indicated in the GitHub dataset available which ones were filled in using this procedure. It matters greatly to how valuable this dataset becomes.**

We have added to the text the number of species with values calculated from species-values, genus-averages or family-averages. We also added this information to the primary data available at the [GitHub repository](https://github.com/efernandezpascual/alpineseeds). We have indicated, in the relevant places of the text, which datasets ([germination](https://github.com/efernandezpascual/alpineseeds/blob/master/data/germination.csv), [species traits](https://github.com/efernandezpascual/alpineseeds/blob/master/data/traits.csv), and [phylogenetic tree](https://github.com/efernandezpascual/alpineseeds/blob/master/data/alpine.tree.pdf)) and results (detailed results of the [MCMCglmm analysis](https://github.com/efernandezpascual/alpineseeds/tree/master/results/MCMCglmm)) are available at GitHub.

> **3) Line 312: Is the phylogeny also available on Github? I’m guessing users would have to run the code to be able to generate it and it would be better if a visual was provided in the supplementary information of the paper itself. **

The phylogeny is available at the [GitHub repository](https://github.com/efernandezpascual/alpineseeds) as a [.tree file](https://github.com/efernandezpascual/alpineseeds/blob/master/data/ALPINE.tree) that can be imported into R, or opened with software like [TreeGraph](http://treegraph.bioinfweb.info/). We have now also included a [pdf version of the tree](https://github.com/efernandezpascual/alpineseeds/blob/master/data/alpine.tree.pdf), so it can be visualized directly. We have indicated at the appropriate point of the text that the tree can be accessed at GitHub. We would prefer this solution to including it as a supplementary file, to avoid duplicating data sources.

> **4) Lines 503-508: This is a very long sentence. Please divide into 2.**

We have divided the sentence in two, as requested.

> **5) Line 610: I think these are the sort of vague statements that Reviewer #2 was referring to regarding how phylogeny is confounding your models and the manuscript could still use some improvement here. For example, you provide phylogenetic signal in variables such as germination time on temperature (although I would argue that Figure 4 makes the values too difficult to distinguish in the 6 panel graphs), but I do not see values for the phylogenetic signal in whether a species is alpine vs. generalist, embryo-seed size, seed mass, etc. Please provide the phylogenetic signal values for all of these traits. Doing so will make statements regarding how “traits are connected and may be subjected to co-adaptation” much easier to interpret. Some improvements in Figure 4 would also be welcome (or perhaps the values can be provided instead in a table).**

We have changed the sentence in former L610 to clarify that we are referring to the phylogenetic signal we found in the germination responses.

Our study focuses on germination responses, and the MCMC models provide a value of phylogenetic signal in (i) the responses themselves [i.e. the main effects], and (2) the interaction between the responses and the other traits (i.e. the moderator effect of seed mass, embryo, strict/generalist). In Figure 4, we only show the signal in the main effects, because the signal in the interactions is almost identical to that in the corresponding main effects. However, the detailed values for each main effect and interaction can be found in the [MCMCglmm results](https://github.com/efernandezpascual/alpineseeds/tree/master/results/MCMCglmm) folder at GitHub (we now specify this in the text).

We have calculated phylogenetic signal in the other traits. It was strong in both seed mass (lambda = 0.94) and the embryo:seed ratio (lambda = 0.99), corresponding to what can be expected from the literature. We also calculated phylogenetic signal in the strict alpine vs. generalist character using D [@RN4959], a metric for binary traits. The value of D was close to 1 (D = 0.92) indicating that the alpine vs. generalist character was randomly distributed across the phylogeny, this is, it had low phylogenetic signal. Nevertheless, we would rather not incorporate these values into the text, for two reasons. First, phylogenetic signal in other traits is not part of the story in this manuscript (rather, our focus would be in the interaction between other traits and germination responses). Second, since seed mass and embryo have been estimated from genus or family values for part of the species, estimates of phylogenetic signal in these traits can be misleading. There are many references focusing more explicitly in phylogenetic signal in seed mass and embryo size, and we discuss some of these in the discussion.

We have revised fonts and sizes of all figures to improve clarity.

> **6) Line 642: As Reviewer #2 mentioned with the previous submission, there is a tendency to state things in a vague way. Here again you mention that tropical systems are subjected to “different selective pressures”. Different how?**

We have revised the entire paragraph, to make clear that we referred to the environmental conditions that we had described in previous sentences: absence of yearly climatic seasonality, strong night-day differences.

# References