### 2020-11-12 README.Rmd
github ms 📰

---
title: "README.md"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

[![DOI](https://zenodo.org/badge/254951190.svg)](https://zenodo.org/badge/latestdoi/254951190)

![Picos de Europa National Park, Cantabrian Mountains, Spain](Picos de Europa.jpg)

# The seed germination spectrum of alpine plants: a global meta-analysis

This repository stores all information related to the manuscript *The seed germination spectrum of alpine plants: a global meta-analysis*, including raw data, `R` scripts to process and analyse data, and `Rmarkdown` files to create the manuscript.

## Contents

This repository is organised following the advice of [Wilson et al. 2017](https://doi.org/10.1371/journal.pcbi.1005510) for recording and storing research projects.

The following materials are available in the folders of this repository:

* `data` Primary data, including the [germination](https://github.com/efernandezpascual/alpineseeds/blob/master/data/germination.csv) dataset, the species [traits](https://github.com/efernandezpascual/alpineseeds/blob/master/data/traits.csv) and the phylogenetic [tree](https://github.com/efernandezpascual/alpineseeds/blob/master/data/ALPINE.tree.pdf).
* `doc` Files to create the [manuscript](https://github.com/efernandezpascual/alpineseeds/blob/master/doc/manuscript.md) using `Rmarkdown`.
* `results` Output of the `R` scripts, including [MCMCglmm analysis](https://github.com/efernandezpascual/alpineseeds/tree/master/results/MCMCglmm) outputs.
* `src` Scripts in `R` language used to process data and perform the analyses of the manuscript.

## Citation

Please cite this repository as:

> Fernández-Pascual E, Carta A, Mondoni A, Cavieres L, Rosbakh S, Venn S, Satyanti A, Guja L, Briceño VF, Vandelook F, Mattana E, Saatkamp A, Bu H, Sommerville K, Poschlod P, Liu K, Nicotra A, Jiménez-Alfaro B (2020) The seed germination spectrum of alpine plants: a global meta-analysis. *New Phytologist*.